### 2023-02-22 outputs/paper/paper.qmd
Add files via upload

---
title: "The Post-2007 Declining U.S. Birth Rates imply Significant Societal Shifts in the U.S."
author: 
  - Sagith Kalaichelvam
  - Youngho Kim
  - Joseph Chung
thanks: "Code, data and reproduction package are available at: https://github.com/UtopianYoungChung/Paper2 and reproduction package link to be updated"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: "This paper explores several potential explanatory factors for the steep decline in birth rates in the United States between 2007 and 2020. The decline in U.S. birthrates offers some significant insights into U.S. society. Changes in birth patterns among recent young adult women moving through the childbearing age group can be directly related to the decline in birth rate from a purely statistical standpoint. We provide evidence that the steep drop in birth rates could be attributed to recent young adults' shifting interests. These shifting priorities represent changes in young adults' preferences regarding having children and aspirations for life. Resonating with the original paper, we conclude with a brief discussion of the implications of the declining birth rate in the United States and its impact on society as an entirety."
    toc: true
    number-sections: true
    colorlinks: true
    lot: false
    lof: true
bibliography: references.bib
---

```{r setup, include = FALSE}
#Load libraries
library(gridExtra)
library(cdlTools)
library(RGraphics)
library(tidyverse)
library(palmerpenguins)
library(knitr)
library(ggplot2)
library(haven)
library(dplyr)
options(dplyr.summarise.inform = FALSE)
```

# Introduction

To measure birth rates, the National Center for Health Statistics (NCHS) aggregates data on birth rates for women by age, race, education level, and other factors; the aggregated data for a particular year is the total number of births per 1000 women. Each year NCHS measures birth rates nationwide and analyzes the data to make insights into societal phenomena. For instance, in 1997, NCHS published a paper on how a mother's education level influences the birth rate. In the paper, NCHS asserted that there is a negative correlation between a mother's education level and birth rate. Their study found a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment @citeNCHS.

In this paper, we gathered data from the American Economic Association, National Center for Health Statistics, and the U.S. population data. With the gathered data, we generated graphs and tables to observe the trend in birth rates of different demographics using R [@citeR]. We also uploaded and cleaned data using R [@citeR], tidyr [@citetidyr], dplyr [@citedplyr], tidyverse [@citetidyverse], and haven [@citehaven] packages. Figures and tables were created with knitr [@citeknit], ggplot2 [@citeggplot], kableExtra [@citekable], and dplyr [@citedplyr].

@fig-1 depicts U.S. birth rate trends over the last 40 years. Here, we can see that the trend is fluctuating relatively stable. However, during the post-2007 recession, it takes a steep decline with no sign of returning. We start our journey to explore potential explanations for the decline.

In @sec-first-point, we discuss how decomposing the decline in the birth rate between demographic groups offers us a clearer picture of the topic. As shown in @fig-2 and @fig-3, by dissecting the demographics into age groups and ethnicity groups, we notice that there is a significant amount of decrease among the white ethnicity women between ages 15 and 24, while other ethnic groups are showing a minor decrease.

@sec-second-point, briefly discusses the empirical relationship between birth rates and economics and policy factors. We then argue that other economic, policy or social factors that have changed since 2007 do not contribute much to the decline.

@sec-third-point, presents a potential attributing factor for the declining U.S. birth rate after the 2007 recession. We show strong evidence that suggests the most recent group of young adults is a significant driver of the post-2007 decline in the birth rate. This significance suggests that the societal shifts among young women moving through childbearing age are an ongoing trend.

We conclude the paper by raising a concern that this societal phenomenon imposes some severe societal consequences and, therefore, should be considered at a high level. Most importantly, we need to consider what it means for the nation's well-being when there are no signs of returning from the negative trend in birth rates.

\newpage

# Data {#sec-data}

## Data Source and Collection

The datasets for this paper were obtained from the paper "The Puzzle of Falling US Birth Rates since the Great Recession" [@citepaper], a paper from the Journal of Economic Perspectives [@citewebsite]. The data used by the paper was compiled by various sources such as the NCHS Natality Database [@citenchs], CDC SEER [@citecdcseer], the Current Population Survey Annual Social and Economic Supplement (CPS ASEC) [@citeipums], the New York Federal Reserve [@citefed], and the Integrated Public Use Microdata Series (IPUMS) [@citeipums]. The paper and datasets were published during the year 2022. All of these various sources are providing a lot of useful information which are shown throughout all of the figures within this paper we created. For example, the sources collected lots of raw data such as population of people giving births, race, lifestyle factors, age of parent when giving birth, etc. So, all of these information were collected from these various sources, and written into the paper "The Puzzle of Falling US Birth Rates since the Great Recession" [@citepaper]. Which was then analyzed in our paper with our version of analysis.

## Data Cleaning

Most of data used in this paper were cleaned to get the most accurate information, and remove any unneccessary information to be outputted in this paper. Data was cleaned to remove data of some years, age group, race category, etc. So, that the revised figures can provide the most accurate information according to our analysis, and not have any information that is unrelevant to our analysis. From our dataset we dropped any values that are not related to the specific category, for example values like NA. We have tried to filter our data much as we can for all of the graphs to make sure that it can makes sense.

## Data Modification

We have included mainly all datasets and variables that are very important for our research question. We included variables by race, birth rates, age group, and other important factors. Lots of other variables were included such as total number of births, and birth rates. Both of these allow us to analyze the decline of births over time. We did lots of modification of the variables for this paper, which was mainly done within the age category. Generally, the original paper have age category as one thing from age 15-44. For our figures, we made several figures and chosen a specific age range. We have tried our best to divide range into categories like 15-19, 20-24, 25-34, and 35-44 through out most of the figures. We have also done modification for race by having a figure that shows the data by different race such as Black, White, and Hispanic people. Several figures had important modification to be relevant to our analysis throughout the paper.

## Data Visualization

Most of the figures shows lots of visualization of the data. We can easily expect that the birth rates is declining over time within the United States due to lots of other factors that is impacting the decision of people to have more kids. When it comes to race, there are lots of data showing some useful information to visualize. Overall, black people tends to have lower birth rates than Hispanic, and white people. Hispanic has the highest birth rates, while while people is the second highest. But all races are showing a down trendline that the birth rates are going down over time which can be coming from other factors.When it comes to factors like child care expenses, it shows that people with lower expense tends to have low birth rates, while more expense tends to have more birth rates. When the rental housing costs is high, the number of birth rates tends to go down, the same for the situation in student loan debt. When it comes into the figures of different races, we decided to focus on age groups such as 15-19, and 20-24 because we believe that this age category has more useful information since we know that in the past, people tend to have kids much earlier than before due to having marriage at an early age. The birth rates by different races in the age group 15-19 has drastically decrease over time when comparing to the age group 20-24. This is because less people are getting marry before age 20, and having kids before age 20.

\newpage

# Results {#sec-results}

Figure 1. Trends in US birth rates over the past 40 years show a steep decline in the birth rate during the post-2007 recession (@fig-1).

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-1
#| fig-cap: Trends in U.S. Birth rates of 1000 women between age 15 to 44 from 1980 to 2020

fig1_data <- read.csv(here::here("inputs/data/fig_1.csv"))
fig1_data %>%
  ggplot(aes(x = year,
             y = brate_all)) +
  geom_line() +
  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
  labs(
    title = "Trend in US Birth Rates",
    x = "Years",
    y = "Births per 1,000 women age 15-44",
    caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019 and 2020."
  )  +
  theme_light()
```
@fig-1 shows that overtime, the number of births in the United States is dropping steadily, while it slightly went up on peak around 1990, 2001, and 2007. Most likely the hikes of going up during these years might have been into other factors that made more people among the selected population to have kids. After 2007, the number of births drastically dropped a lot. Nowadays, among 1,000 women from age 15 to 44, only about 56 births will occur. While, in the past around the 1980s, and 1990s, the number of births was around 65 to 70 range. The line in the graph shows that in the near future there will be much lower amount of people to have kids, and this can be due to other factors from their lifestyle which is making them to decide to not have more kids.

While the figure shows the birth rate trend is going down, there are certain spikes on peak of birth rates around 1990, 2001, and 2007. There were some incidents that must have happened in the United States in those years which may have caused people to stay home more often. As we know that people that are likely to be unemployed tends to have more kids as they will more freedom during their time spent home. This allows the birth rates to be higher than usual. During these years, the Americans were facing lower opportunities in employment, and within the society [@citecatholic]. They do not have access to useful resources that improve their lifestyle in these years, so they are at home at most of the times, which encourages couples to have an interest of having more kids, which ends up increasing the birth rates for these years.

\newpage

Figure 2.place it here with Figure 6 (@fig-2).

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-2
#| fig-cap: Race and Ethnicity (ages 15-19 from 2000 to 2020)
birth1 = read_dta("data/nchs/nchs_births_pop_1990_2019.dta")

birth1 <- birth1 %>%
  group_by(year) %>%
  select(-c(stname)) %>% 
  summarise(across(everything(),sum))

# getting total number of births for white people from 15 to 19 years old
birth1$numbirthwhite15 = birth1$numbirth_lesshs_whitenh_1519 + birth1$numbirth_educ_miss_whitenh_1519 + birth1$numbirth_gtc_whitenh_1519 + birth1$numbirth_somecoll_whitenh_1519 + birth1$numbirth_hsgrad_whitenh_1519 + birth1$numbirth_ba_whitenh_1519

# getting total number of births for black people from 15 to 19 years old  
birth1$numbirthblack15 = birth1$numbirth_lesshs_blacknh_1519 + birth1$numbirth_educ_miss_blacknh_1519 + birth1$numbirth_gtc_blacknh_1519 + birth1$numbirth_somecoll_blacknh_1519 + birth1$numbirth_hsgrad_blacknh_1519 + birth1$numbirth_ba_blacknh_1519

# getting total number of births for hispanics from 15 to 19 years old  
birth1$numbirthhisp15 = birth1$numbirth_lesshs_hisp_1519 + birth1$numbirth_educ_miss_hisp_1519 + birth1$numbirth_gtc_hisp_1519 + birth1$numbirth_somecoll_hisp_1519 + birth1$numbirth_hsgrad_hisp_1519 + birth1$numbirth_ba_hisp_1519

# getting total number of births for white people from 20 to 24 years old
birth1$numbirthwhite20 = birth1$numbirth_lesshs_whitenh_2024 + birth1$numbirth_educ_miss_whitenh_2024 + birth1$numbirth_gtc_whitenh_2024 + birth1$numbirth_somecoll_whitenh_2024 + birth1$numbirth_hsgrad_whitenh_2024 + birth1$numbirth_ba_whitenh_2024

# getting total number of births for black people from 20 to 24 years old
birth1$numbirthblack20 = birth1$numbirth_lesshs_blacknh_2024 + birth1$numbirth_educ_miss_blacknh_2024 + birth1$numbirth_gtc_blacknh_2024 + birth1$numbirth_somecoll_blacknh_2024 + birth1$numbirth_hsgrad_blacknh_2024 + birth1$numbirth_ba_blacknh_2024

# getting total number of births for hispanics from 20 to 24 years old  
birth1$numbirthhisp20 = birth1$numbirth_lesshs_hisp_2024 + birth1$numbirth_educ_miss_hisp_2024 + birth1$numbirth_gtc_hisp_2024 + birth1$numbirth_somecoll_hisp_2024 + birth1$numbirth_hsgrad_hisp_2024 + birth1$numbirth_ba_hisp_2024

# gets the total number of births by year, compiled nationally
birthhist = birth1 %>%
  filter(year >= 2000) %>%
  group_by(year) %>% 
  select(year,pop1544, numbirthwhite15,numbirthwhite20,numbirthhisp15,numbirthhisp20,numbirthblack15,numbirthblack20) %>%
  summarize(sumpop = sum(pop1544),white15 = sum(numbirthwhite15), white20 = sum(numbirthwhite20), black15 = sum(numbirthblack15), black20 = sum(numbirthblack20), hisp15 = sum(numbirthhisp15), hisp20 = sum(numbirthhisp20))

# gets the birth rate for each race by adding the number of births from the 15 to 19for each race, then multiplying by 10,000 and dividing by the female population
birthhist %>%
  mutate(whiteall = white15*10000/sumpop, blackall= black15*10000/sumpop, hispall = hisp15*10000/sumpop) %>% 
  select(year, whiteall, blackall, hispall) %>%
  pivot_longer(cols = c("whiteall", "blackall", "hispall"),
               names_to = "racegroup",
               values_to = "value"
  ) %>%
  ggplot(mapping = aes(x = year, y = value, color = racegroup)) +
  geom_line() +
  theme_minimal() +
  scale_x_continuous(breaks = seq(2000,2020, by=2))+
  labs(title = "birthrate for each race in age group 15 to 19",
       x = "Year",
       y = "Birth Rate",
       color = "Race") +
  scale_color_brewer(palette = "Set1", labels = c("Black (Non-Hispanic)", "Hispanic", "White (Non-Hispanic)")) +
  theme(legend.position = "bottom")

```

\newpage

Figure 3. place it here with Figure 7 (@fig-3).

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-3
#| fig-cap: Race and Ethnicity (ages 20-24 from 2000 to 2020)

birth1 = read_dta("data/nchs/nchs_births_pop_1990_2019.dta")

birth1 <- birth1 %>%
  group_by(year) %>%
  select(-c(stname)) %>% 
  summarise(across(everything(),sum))

# getting total number of births for white people from 15 to 19 years old
birth1$numbirthwhite15 = birth1$numbirth_lesshs_whitenh_1519 + birth1$numbirth_educ_miss_whitenh_1519 + birth1$numbirth_gtc_whitenh_1519 + birth1$numbirth_somecoll_whitenh_1519 + birth1$numbirth_hsgrad_whitenh_1519 + birth1$numbirth_ba_whitenh_1519

# getting total number of births for black people from 15 to 19 years old  
birth1$numbirthblack15 = birth1$numbirth_lesshs_blacknh_1519 + birth1$numbirth_educ_miss_blacknh_1519 + birth1$numbirth_gtc_blacknh_1519 + birth1$numbirth_somecoll_blacknh_1519 + birth1$numbirth_hsgrad_blacknh_1519 + birth1$numbirth_ba_blacknh_1519

# getting total number of births for hispanics from 15 to 19 years old  
birth1$numbirthhisp15 = birth1$numbirth_lesshs_hisp_1519 + birth1$numbirth_educ_miss_hisp_1519 + birth1$numbirth_gtc_hisp_1519 + birth1$numbirth_somecoll_hisp_1519 + birth1$numbirth_hsgrad_hisp_1519 + birth1$numbirth_ba_hisp_1519

# getting total number of births for white people from 20 to 24 years old
birth1$numbirthwhite20 = birth1$numbirth_lesshs_whitenh_2024 + birth1$numbirth_educ_miss_whitenh_2024 + birth1$numbirth_gtc_whitenh_2024 + birth1$numbirth_somecoll_whitenh_2024 + birth1$numbirth_hsgrad_whitenh_2024 + birth1$numbirth_ba_whitenh_2024

# getting total number of births for black people from 20 to 24 years old
birth1$numbirthblack20 = birth1$numbirth_lesshs_blacknh_2024 + birth1$numbirth_educ_miss_blacknh_2024 + birth1$numbirth_gtc_blacknh_2024 + birth1$numbirth_somecoll_blacknh_2024 + birth1$numbirth_hsgrad_blacknh_2024 + birth1$numbirth_ba_blacknh_2024

# getting total number of births for hispanics from 20 to 24 years old  
birth1$numbirthhisp20 = birth1$numbirth_lesshs_hisp_2024 + birth1$numbirth_educ_miss_hisp_2024 + birth1$numbirth_gtc_hisp_2024 + birth1$numbirth_somecoll_hisp_2024 + birth1$numbirth_hsgrad_hisp_2024 + birth1$numbirth_ba_hisp_2024

# gets the total number of births by year, compiled nationally
birthhist = birth1 %>%
  filter(year >= 2000) %>%
  group_by(year) %>% 
  select(year,pop1544, numbirthwhite15,numbirthwhite20,numbirthhisp15,numbirthhisp20,numbirthblack15,numbirthblack20) %>%
  summarize(sumpop = sum(pop1544),white15 = sum(numbirthwhite15), white20 = sum(numbirthwhite20), black15 = sum(numbirthblack15), black20 = sum(numbirthblack20), hisp15 = sum(numbirthhisp15), hisp20 = sum(numbirthhisp20))

# gets the birth rate for each race by adding the number of births from the 15 to 19for each race, then multiplying by 10,000 and dividing by the female population
birthhist %>%
  mutate(whiteall = white20*10000/sumpop, blackall= black20*10000/sumpop, hispall = hisp20*10000/sumpop) %>% 
  select(year, whiteall, blackall, hispall) %>%
  pivot_longer(cols = c("whiteall", "blackall", "hispall"),
               names_to = "racegroup",
               values_to = "value"
  ) %>%
  ggplot(mapping = aes(x = year, y = value, color = racegroup)) +
  geom_line() +
  theme_minimal() +
  scale_x_continuous(breaks = seq(2000,2020, by=2))+
  labs(title="birthrate for each race in age group 20 to 24",
       x = "Year",
       y = "Birth Rate",
       color = "Race") +
  scale_color_brewer(palette = "Set1", labels = c("Black (Non-Hispanic)", "Hispanic", "White (Non-Hispanic)")) +
  theme(legend.position = "bottom")

```

\newpage

Figure 4. Relationship between changes in birth rates and potential explanatory factors (@fig-4).
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-4
#| fig-cap: Other Economic factors over the periods of 2004-2008 and 2014-2019

#04~19 birthrate data
birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))
pop1 = read_dta(here::here("inputs/data/pop/age_race_comp_seer.dta"))
birth1 <- birth1 %>% 
  filter((year>=2004 & year<=2008) | (year>=2015 & year<=2019))

birth1 <- birth1 %>% 
  subset(select=c(stname,year,numbirth1544))

pop1<-pop1 %>% 
  mutate(pop1544 = pop1519 + pop2034 + pop3544)%>% 
  filter((year>=2004 & year<=2008) | (year>=2015 & year<=2019)) %>%
  group_by(year,stname) %>%
  select(pop1544) %>% 
  summarise(across(everything(),sum))

fig4_data <- right_join(birth1, pop1, by=c("stname","year"))
fig4_data <- fig4_data %>% 
  mutate(year = case_when(year=2004 & year<=2008 ~ 2004,
                           year>=2015 & year<=2019 ~ 2019)) %>%
  group_by(year,stname) %>%
  select(numbirth1544,pop1544) %>% 
  summarise(across(everything(),sum)) %>% 
  mutate(birthrate=numbirth1544/pop1544*1000) %>% 
  select(-c(numbirth1544,pop1544))

fig4_data <- fig4_data %>% 
  group_by(stname) %>% 
  mutate(birthrate = birthrate[year == 2019] - birthrate[year == 2004])

#04~19 childcare cost for fig4b
childcare_aesc=read_dta(here::here("inputs/data/long_term/childcare_asec_09_1519.dta"))
childcare_aesc <- childcare_aesc %>% 
  mutate(childcare_cost=child_asec)
childcare_aesc <- childcare_aesc %>% 
  select(year,stname,childcare_cost)
fig4_data <- right_join(fig4_data, childcare_aesc, by=c("stname","year"))
fig4_data <- fig4_data %>% 
  group_by(stname) %>% 
  mutate(childcare_cost = childcare_cost[year == 2019] - childcare_cost[year == 2004])

#04~19 rentcost data for fig4c
rentcost_0419=read_dta(here::here("inputs/data/long_term/longdiff-rentcost-0419.dta"))

rentcost_0419 <- rentcost_0419 %>% 
  mutate(stname = fips(rentcost_0419$statefip, to='Abbreviation')) %>%
  group_by(stname) %>% 
  mutate(avg_grossrent = avg_grossrent[year == 2019] - avg_grossrent[year == 2004]) %>% 
  select(year,stname,avg_grossrent)

fig4_data <- right_join(fig4_data, rentcost_0419, by=c("stname","year"))

#04~19 wage ratio data for fig4d
rhs_1519=read_dta(here::here("inputs/data/long_term/longdiff-RHS-1519.dta"))
rhs_0408=read_dta(here::here("inputs/data/long_term/longdiff-RHS-0408.dta"))
rhs<-rbind(rhs_1519, rhs_0408)
rhs <- rhs %>% 
  mutate(stname = fips(rhs$statefip, to='Abbreviation')) %>% 
  mutate(fem_male_ratio=p50wage_all/p50wage_male_all) %>% 
  select(year,stname,fem_male_ratio) %>% 
  group_by(stname) %>% 
  mutate(fem_male_ratio = fem_male_ratio[year == 2019] - fem_male_ratio[year == 2004])

fig4_data <- right_join(fig4_data, rhs, by=c("stname","year"))

#04~19 student loan data for fig4e
std_loan=read_dta(here::here("inputs/data/long_term/nyfed-student-debt-per-capita-0419.dta"))

std_loan<- std_loan %>% 
  group_by(stname) %>% 
  mutate(stud_loan_cap = stud_loan_cap[year == 2019] - stud_loan_cap[year == 2004])

fig4_data <- right_join(fig4_data, std_loan, by=c("stname","year"))

fig4_data <- fig4_data %>% 
  head(51) %>% 
  select(-c(year))

#drawing
fig4b<-fig4_data %>% 
  ggplot(aes(x=childcare_cost*1000, y=birthrate)) +
  geom_point(size=0)+
  labs(
    title = "Change in childcare expenditures",
    x = "US dollar",
    y = "birthrate") +
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(0, 5000), breaks=seq(0, 5000, by = 1000),labels = paste0("$",seq(0, 5000, by=1000))) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

fig4c <-fig4_data %>% 
  ggplot(aes(x=avg_grossrent, y=birthrate)) +
  geom_point(size=0)  +
  labs(
    title = "Change in rental housing costs",
    x = "US dollar",
    y = "birthrate") +
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(0, 400), breaks=seq(0, 400, by = 75),labels = paste0("$",seq(0, 400, by = 75))) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

fig4d<-fig4_data %>% 
  ggplot(aes(x=fem_male_ratio, y=birthrate)) +
  geom_point(size=0)  +
  labs(
    title = "Change in female-male wage ratio",
    x = "fem to male wage ratio",
    y = "birthrate")+
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(-0.08, 0.12), breaks=seq(-0.08, 0.12, by = 0.04)) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

fig4e<-fig4_data %>% 
  ggplot(aes(x=stud_loan_cap*1000, y=birthrate)) +
  geom_point(size=0)  +
  labs(
    title = "Change in per capita student loan debt",
    x = "US dollar",
    y = "birthrate") +
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(1000, 4500), breaks=seq(1000, 4500, by = 750),labels = paste0("$",seq(1000, 4500, by=750))) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

grid.arrange(fig4b,fig4c,fig4d,fig4e,ncol=2,nrow=2)

```

@fig-4 shows interesting relationship between changes in birthrate and other economical factors over five years period. Each marker represents a state, labeled with its two-letter abbreviation.

Annual expenditures on childcare for families with children under 12 whoreport positive childcare spending rose $2,170 nationwide for the average of 2015–2019, based on data from the Current Population Survey (all dollar values throughout are measured in constant 2019 dollars). But in the cross-state correlations, places where childcare expenditures increased more did not experience a noticeable drop in birth rates. In fact, the trend shows that more increases of childcare expenditure has no significant impact on birthrate changes.

Average monthly rents for a two- to three-bedroom apartment rose $124 per month (from $930 to $1,060 in 2018 dollars, a 14 percent increase) nationwide over this period. The increase was much larger in some states, like Colorado, Washington, and the District of Columbia. The data, however, do not indicate a negative association between changes in rents and changes in birth rates.

We used the female-to-male median earnings ratio among full-time, year-round workers to look at women's economic opportunity. On average, the midian earning ratio rose 0.04 in the period. 12 States in which the wage ratio rose the most over this sample period did not exhibit a greater decline in the birth rate.

If young adults are saddled with debt, they might not feel like they have sufficient disposable income to have a child or more children. We consider the total level of student debt per capita in a state, which has increased from $2,500 to $5,400 (in 2018 dollars), on average. The relationship between state-level student debt and the birth rate are generally flat. It indicates that increases in student debt are related to the aggregate reductions in birth rates.

\newpage

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-5
#| fig-cap: Relationship between Changes in Birth Rates 2004-2008 to 2014-2019

# Reading in the data.

birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))

birth1 <- birth1 %>%
  group_by(year) %>%
  select(-c(stname)) %>% 
  summarise(across(everything(),sum))

birth1 <- birth1 %>% 
  mutate(birthrate1519 = numbirth1519/pop1519*1000) %>% 
  mutate(birthrate2024 = numbirth2024/pop2024*1000) %>% 
  mutate(birthrate2529 = numbirth2529/(pop25+pop26+pop27+pop28+pop29)*1000) %>% 
  mutate(birthrate3034 = numbirth3034/(pop30+pop31+pop32+pop33+pop34)*1000) %>% 
  mutate(birthrate3539 = numbirth3539/(pop35+pop36+pop37+pop38+pop39)*1000) %>% 
  mutate(birthrate4044 = numbirth4044/(pop40+pop41+pop42+pop43+pop44)*1000)

birth1 <- birth1 %>% 
  select(year, birthrate1519, birthrate2024, birthrate2529, birthrate3034, birthrate3539, birthrate4044) %>%
  pivot_longer(-c(year),
               names_to = "agegroup",
               values_to = "birthrate"
  )

birth1<-birth1 %>% 
  mutate(agegroup = as.character(str_replace(agegroup, "birthrate", "")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "1519", "15~19")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "2024", "20~24")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "2529", "25~29")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "3034", "30~34")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "3539", "35~39")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "4044", "40~44")))

birth1 %>% 
ggplot(mapping = aes(x = year, y = birthrate, color = agegroup)) + 
  geom_line()  +
  labs(
    title = "Trends in Birth Rates by Age Group",
    x = "Year",
    y = "birthrate",
    caption = "Source: CDC Vital Statistics Births Reports.",
    color = "Age Group"
  ) +
  scale_x_continuous(limits=c(1991, 2020), breaks=seq(1991, 2020, by = 8)) +
  scale_y_continuous(limits = c(0,125), breaks = seq(0,125, by=20)) +
  theme_light()
```

@fig-5 shows lots of important information about when people tend to have kids at certain age over the timeframe from 1990 to 2020. Based on the graph, the data shows that more people are having kids after the age 30 during this present time. In the past, most people tend to have kids before the age 30. Around the years 2000 to 2005, that is when the lifestyle started to change where people tend to either get marry later, or have kids later which is why the trend of birth rates in age 30 and up are going higher from the 2000's. While, the births before the age of 30 tends to slow down and decreasing over time from the 2000's. According to the graph, maybe in another 100 years, we can expect that most birth rates can happen around the age 40-44.

There are drastic changes shown in @fig-5 such as age 30-44 tends to increase around the 2000s, while the ages below 30 decreases for amount of birth rates. Around the decade of 2000's or beginning of this new 21st century, there is lots of modernization from people's lifestyle to other things in Earth which is coming from several factors such as education and employment due to new innnovation and invention [@citedimock]. There are lots of new behaviour being introduced to new generation of people. So, as they get new behaviour of attiudes, that is being involved in their decisions of marriages, and when to have kids. This is why lots of people are now getting married late and/or having kids much later while comparing to the past.

\newpage

Figure 6. Children Ever Born by Mother's Age and Birth Cohort (@fig-6).

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-6
#| fig-cap: Children Ever Born by Mother's Age and Birth Cohort

nch_cohot_data <- read_dta(here::here("inputs/data/nchs/nchs_cohort_analysis.dta"))
agecomp_data <- read_dta(here::here("inputs/data/pop/agecomp-seer.dta"))

pop_by_cohort<- agecomp_data %>%
  select(1:32) %>%
  group_by(year) %>%
  select(-c(stname)) %>%
  summarise(across(everything(),sum)) %>%
  pivot_longer(-c(year), names_to = "mage", values_to = "pop") %>%
  mutate(mage = as.numeric(str_replace(mage, "fem", ""))) %>%
  mutate(birthyear = year - mage) %>%
  mutate(cohort2 = case_when(birthyear >= 1968 & birthyear <=1972 ~ 1,
                            birthyear >= 1973 & birthyear <=1977 ~ 2,
                            birthyear >= 1978 & birthyear <=1982 ~ 3,
                            birthyear >= 1983 & birthyear <=1987 ~ 4,
                            birthyear >= 1988 & birthyear <=1992 ~ 5,
                            birthyear >= 1993 & birthyear <=1997 ~ 6)) %>% 
  group_by(cohort2,mage) %>%
  summarise(pop = sum(pop))

fig5_data <- right_join(nch_cohot_data, pop_by_cohort, by=c("mage", "cohort2"))
fig5_data <- fig5_data %>%
  mutate(birthrate=numbirth/pop*1000,
         age_2024_year = case_when(cohort2 == 1 ~ 1992,
                                   cohort2 == 2 ~ 1997,
                                   cohort2 == 3 ~ 2002,
                                   cohort2 == 4 ~ 2007,
                                   cohort2 == 5 ~ 2012,
                                   cohort2 == 6 ~ 2017))

fig5_data <- fig5_data %>% 
  group_by(age_2024_year) %>% 
  mutate(cum_birthrate = cumsum(birthrate)/1000, 
         cohort = case_when(cohort2 == 1 ~ "1968-1972        1992",
                                  cohort2 == 2 ~ "1973-1977        1997",
                                  cohort2 == 3 ~ "1978-1982        2002",
                                  cohort2 == 4 ~ "1983-1987        2007",
                                  cohort2 == 5 ~ "1988-1992        2012",
                                  cohort2 == 6 ~ "1993-1997        2017"))

fig5_data <-  na.omit(fig5_data)
fig5_data %>%
  ggplot(aes(x = mage,
             y = cum_birthrate,
             color = cohort)) +
  geom_line()  +
  labs(
    title = "Children Ever Born by Mother’s Age and Birth Cohort",
    x = "Ages",
    y = "Cumulative birthrate",
    caption = "Source: NCHS Vital Statistics birth microdata and CDC SEER population data.",
    color = "Birth cohorts   Age 20 in"
  ) +
  scale_x_continuous(limits=c(15, 44), breaks=seq(15, 44, by = 3)) +
  theme_minimal()+
  theme(legend.position = "right")

```
@fig-6 shows lots of useful information about the cumulative birth rate by mother's age in several birth cohorts from early as 1968 to late as 1997. We tend to use these years since in recent times we believe that there is enough data for these years since most of the births and living people are among these cohorts. According to @fig-6, the earlier the cohort is, then the cumulative birth rate will be really high. While, the later the cohort is, then the cumulative birth rate will be much lower. The highest cumulative birth rate is coming from the birth cohort of 1968-1972, and 1973-1977. While the lowest is coming from the 1993-1997 birth cohort.

\newpage

# Discussion

## Birth Rates by Demographic Groups {#sec-first-point}
In 1997, NCHS released a report titled "Mother's Educational Level Influences Birth Rate." According to the report, a women's education level is the best predictor of how many children she will have. In the report, NCHS describes a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment [@citeNCHS].  
We are particularly interested in ethnic groups of ages 15-19 and 20-24. @fig-2 shows a sharp decline in birth rates among all groups of women between the age of 15 to 19 during the post-2007 recession. Figure 3 also shows a sharp decline over the same period, however, only among white ethnic women between the ages 20 to 24, while other groups show a relatively steady decline. Figures that are omitted from this paper of the other age groups (25 to 34 and 35 to 44) show a similar trend both in pre and post-recession.
However, as shown in table 1, the most significant contributor to the decline in the birth rate is Hispanic, with less than High School education, women between ages 20-24 after 2007. This result contradicts the earlier claim made by NCHS. The finding suggests, in the post-2007 recession, the women's education level and birth rate now correlate positively.

## Economic and Policy Factors {#sec-second-point}
In @fig-4, we show other economic factors that might have negatively affected the birth rate in the post-2007 recession across the states. Although the graphs show their inverse relationship, no dramatic change is shown. For instance, as there were somewhere around $1000 to $2000 increases in childcare expenses over the past 13 years, we see that the decrease in the actual birth rate is only between 0 to -10. Change in per capita student loans also has increased between $2500 to $3250, but it only contributed to the decrease somewhere between 0 to -5 across the states. Based on these findings, it is hard to find an explanation of the dramatic decline in the birth rate since 2007 from economic factors.

## A Potential Explanation: Shifting Priorities {#sec-third-point}

When it comes to the age of all the population from the United States and comparing it from the past to now. It shows that more people tend to have kids much later than before. While, less people have kids earlier than the past. During the past, more people tend to have kids much earlier. There are several reasons why more people had kids at early age because back then people use to get marry much earlier, and many people were unemployed or doing a job for self-employment, which was not even an educated or professional job. The modern lifestyle that is being adapted in the past few decades tends to make people to focus on all other areas of life, which makes them to marry much later, and have kids late. Nowadays, many marriages happen at the age of late 20s or early 30s, while they tend to have their first child around early to mid 30s.
@fig-5 depicts the birth rates in women between the ages 30-39 and 35-39 positive slope trend. Whether pre or post-2007 recession, this trend looks steady over the past 40 years. This finding aligns with the current common belief and the social norm where, perhaps, women's peak childbearing ages have shifted from the early twenties to early thirties. @fig-6 shows the average number of children born to women by age in five-year intervals from the 1968-1972 through 1993-1997 birth cohorts. These birth cohorts include women entering their peak childbearing years (20–24) in 1992, 1997, 2002, 2007, and 2017. This chart indicates that the three cohorts of young adult women born between 1968 and 1982 in 1992, 1997, and 2002 had similar childbearing age profiles. The 2007 birth cohort of women had fewer children in their 20s and 30s. The 2012 and 2017 birth cohorts (born between 1988 and 1997) have significantly fewer children than the preceding cohorts. This finding is consistent with a decline in births, primarily to more recent cohorts of women. From a statistical point of view, this divergence across cohorts explains the sizable decrease in annual birth rates that began in 2007.

## Weakness and Moving Forward

### Bias and Ethical Concerns

As we know that this data was collected and created by other human beings just as how we are. There are some ethical concerns that could affect the accuracy of our data because humans can make mistakes, they can make decisions based on their preferences, which can affect the whole dataset. So, this analysis is only valid based on our dataset, which can be modified if the original dataset was changed.

According to the original data provided which shows different races such as White, Black, and Hispanic. They have included Hispanic based on all colours, which kind of erases the identity of Black and Indigenous Hispanic people. This tends to be a bit racism because the information provided on the data did not specifically indicate whether or not the Hispanic category includes everyone that identifies as Hispanic, or is it just people that look like Hispanic.

### Data Collection Concerns

This dataset tends to collect a huge data information of population among the people from the United States, and considers lots of factors from race, birth rates by age, student debts, etc. Having too much information mean that when the data was collected there is a possibility of human error, calculation error, and chances of small modification of data due format of reporting error. So, the way how the data was collected can have lots of problem because people can make mistakes easily.

### Moving Forward
This paper raised concerns about a dramatic decline in U.S. birth rates starting in 2007. Through the paper, we proposed a general explanation for the decline in births across recent cohorts of U.S. women that focuses on the shifting priorities of cohorts. As dictated by the recent socio-phenomena that encompasses preferences for having children, life aspirations, and the nature of parenting, a change in the relationship between education and childbearing, a rise in childlessness, and the establishment of a two-child norm for those having children have irreversibly changed [@citepaper].
This new social phenomenon that might be causing the dramatic decline in U.S. birth rates is important from an economic policy standpoint to acknowledge that an aging population and shrinking workforce pose challenges to economic growth and the sustainability of social insurance systems [@Yang2003]. Thus, the most appropriate way to address declining U.S. birth rates may be required: 1. developing A.I. and robots where a shrinking workforce poses a problem, 2. an appropriate policy to address issues with Social Security and Medicare on a secure basis for the long-term. Suppose the U.S. birth rate is not showing any reverse trend shortly. In that case, the U.S. economy and political system should enforce greater emphasis on addressing the issue before it becomes irreversible

\newpage

# References