### 2023-02-23 outputs/paper/paper.qmd
Add files via upload

---
title: "The Post-2007 Declining U.S. Birth Rates imply Significant Societal Shifts in the U.S."
author: 
  - Sagith Kalaichelvam
  - Youngho Kim
  - Joseph Chung
thanks: "Code, data and reproduction package are available at: https://github.com/UtopianYoungChung/Paper2 and https://doi.org/10.1257/jep.36.1.151"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: "This paper explores several potential explanatory factors for the steep decline in birth rates in the United States between 2007 and 2020. The decline in U.S. birthrates offers some significant insights into U.S. society. Changes in birth patterns among recent young adult women moving through the childbearing age group can be directly related to the decline in birth rate from a purely statistical standpoint. We provide evidence that the steep drop in birth rates could be attributed to recent young adults' shifting interests. These shifting priorities represent changes in young adults' preferences regarding having children and aspirations for life. Resonating with the original paper, we conclude with a brief discussion of the implications of the declining birth rate in the United States and its impact on society as an entirety."
    toc: true
    number-sections: true
    colorlinks: true
    lot: true
    lof: true
bibliography: references.bib
---

```{r setup, include = FALSE}
#Load libraries
library(gridExtra)
library(cdlTools)
library(RGraphics)
library(tidyverse)
library(palmerpenguins)
library(knitr)
library(ggplot2)
library(haven)
library(dplyr)
options(dplyr.summarise.inform = FALSE)
```

# Introduction

To measure birth rates, the National Center for Health Statistics (NCHS) aggregates data on birth rates for women by age, race, education level, and other factors; the aggregated data for a particular year is the total number of births per 1000 women. Each year NCHS measures birth rates nationwide and analyzes the data to make insights into societal phenomena. For instance, in 1997, NCHS published a paper on how a mother's education level influences the birth rate. In the paper, NCHS asserted that there is a negative correlation between a mother's education level and birth rate. Their study found a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment @citeNCHS.

In this paper, we gathered data from the American Economic Association, National Center for Health Statistics, and the U.S. population data. With the gathered data, we generated graphs and tables to observe the trend in birth rates of different demographics using R [@citeR]. We also uploaded and cleaned data using R [@citeR], tidyr [@citetidyr], dplyr [@citedplyr], tidyverse [@citetidyverse], and haven [@citehaven] packages. Figures and tables were created with knitr [@citeknit], ggplot2 [@citeggplot], kableExtra [@citekable], and dplyr [@citedplyr].

@fig-1 depicts U.S. birth rate trends over the last 40 years. Here, we can see that the trend is fluctuating relatively stable. However, during the post-2007 recession, it takes a steep decline with no sign of returning. We start our journey to explore potential explanations for the decline.

In @sec-first-point, we discuss how decomposing the decline in the birth rate between demographic groups offers us a clearer picture of the topic. As shown in @fig-2, by dissecting the demographics into age groups and ethnicity groups, we notice a significant decrease among the white ethnicity women between ages 15 and 24, while other ethnic groups are showing a minor decrease.

@sec-second-point, briefly discusses the empirical relationship between birth rates and economics and policy factors. We then argue that other economic, policy or social factors that have changed since 2007 do not contribute much to the decline.

@sec-third-point, presents a potential attributing factor for the declining U.S. birth rate after the 2007 recession. We show strong evidence that suggests the most recent group of young adults is a significant driver of the post-2007 decline in the birth rate. This significance suggests that the societal shifts among young women moving through childbearing age are an ongoing trend.

We conclude the paper by raising a concern that this societal phenomenon imposes some severe societal consequences and, therefore, should be considered at a high level. Most importantly, we need to consider what it means for the nation's well-being when there are no signs of returning from the negative trend in birth rates.

# Data {#sec-data}

## Data Source and Collection

The datasets for this paper were obtained from the paper "The Puzzle of Falling US Birth Rates since the Great Recession" [@citepaper], a paper from the Journal of Economic Perspectives [@citewebsite]. The data used by the paper was compiled from various sources such as the NCHS Natality Database [@citenchs], CDC SEER [@citecdcseer], the Current Population Survey Annual Social and Economic Supplement (CPS ASEC) [@citeipums], the New York Federal Reserve [@citefed], and the Integrated Public Use Microdata Series (IPUMS) [@citeipums]. The paper and datasets were published in the year 2022. These various sources provide a lot of useful information, shown throughout all of the figures we created in this paper. For example, the sources collected lots of raw data such as the population of people giving birth, race, lifestyle factors, age of parent when giving birth, etc. So, all of this information was collected from these various sources and written into the paper "The Puzzle of Falling US Birth Rates since the Great Recession" [@citepaper]. Which was then analyzed in our paper with our version of the analysis.

## Data Cleaning

Most of the data used in this paper were cleaned to get the most accurate information and remove any unnecessary information to be outputted in this paper. Data were cleaned to remove data of some years, age groups, race categories, etc. So that the revised figures can provide the most accurate information according to our analysis and not have any information that is unrelevant to our analysis. From our dataset, we dropped any values that are not related to the specific category, for example, values like NA. We have tried to filter our data much as we can for all of the graphs to make sure that it can make sense.

## Data Modification

We have included mainly all datasets and variables that are very important for our research question. We included variables by race, birth rates, age group, and other important factors. Lots of other variables were included, such as the total number of births and birth rates. Both of these allow us to analyze the decline of births over time. We did lots of modifications of the variables for this paper, which were mainly done within the age category. Generally, the original paper has an age category as one thing from age 15-44. For our figures, we made several figures and chose a specific age range. We have tried our best to divide the range into categories like 15-19, 20-24, 25-34, and 35-44 throughout most of the figures. We have also modified for race by having a figure that shows the data by different races, such as Black, White, and Hispanic people. Several figures had an important modifications to be relevant to our analysis throughout the paper.

## Data Visualization

Most of the figures show lots of visualization of the data. We can easily expect that the birth rate is declining over time within the United States due to many other factors impacting people's decision to have more kids. When it comes to race, there are a lot of data showing some useful information to visualize. Overall, black people tend to have lower birth rates than Hispanic and white people. Hispanics have the highest birth rates, while people are the second highest. But all races are showing a down trendline that the birth rates are going down over time which can be coming from other factors. When it comes to factors like childcare expenses, it shows that people with lower expense tends to have low birth rates, while more expense tends to have more birth rates. When the rental housing costs are high, the number of birth rates tends to go down, and the same for the situation in student loan debt. When it comes to the figures of different races, we decided to focus on age groups such as 15-19 and 20-24 because we believe that this age category has more useful information since we know that in the past, people tended to have kids much earlier than before due to having marriage at an early age. The birth rates by different races in the age group 15-19 have drastically decreased over time when compared to the age group 20-24. This is because fewer people are getting married before age 20 and having kids before age 20.

# Results {#sec-results}

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-1
#| fig-cap: Trends in U.S. Birth rates of 1000 women between age 15 to 44 from 1980 to 2020

fig1_data <- read.csv(here::here("inputs/data/fig_1.csv"))
fig1_data %>%
  ggplot(aes(x = year,
             y = brate_all)) +
  geom_line() +
  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
  labs(
    title = "",
    x = "Years",
    y = "Births per 1,000 women age 15-44",
    caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019 and 2020."
  )  +  
  geom_vline(xintercept = 2007, linetype = "dashed") +
  geom_text(aes(x = 2007, y = 70), nudge_x = 2, nudge_y = 1, label = "2007") +
  theme_light()
```

@fig-1 shows that over time, the number of births in the United States is dropping steadily, while it slightly went up on peak around 1990, 2001, and 2007. Most likely, the hikes going up during these years might have been into other factors that made more people among the selected population have kids. After 2007, the number of births drastically dropped a lot. Nowadays, among 1,000 women from age 15 to 44, only about 56 births will occur. While in the past, around the 1980s and 1990s, the number of births was around the 65 to 70 range. The line in the graph shows that in the near future, there will be a much lower amount of people having kids, and this can be due to other factors from their lifestyle, which is making them decide not to have more kids.

While the figure shows the birth rate trend is going down, there are certain spikes in the peak of birth rates around 1990, 2001, and 2007. There were some incidents that must have happened in the United States in those years, which may have caused people to stay home more often. As we know that people that are likely to be unemployed tend to have more kids as they will have more freedom during their time spent at home. This allows the birth rates to be higher than usual. During these years, Americans faced lower employment opportunities within society [@citecatholic]. They do not have access to useful resources that improve their lifestyle in these years, so they are at home most of the time, which encourages couples to have an interest in having more kids, which ends up increasing the birth rates for these years.

\newpage

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-2
#| fig-cap: Trend in Birthrate by Race and Ethnicity for Ages 15-19 and 20-24 from 2000 to 2020)
birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))

birth1 <- birth1 %>%
  group_by(year) %>%
  select(-c(stname)) %>% 
  summarise(across(everything(),sum))

# getting total number of births for white people from 15 to 19 years old
birth1$numbirthwhite15 = birth1$numbirth_lesshs_whitenh_1519 + birth1$numbirth_educ_miss_whitenh_1519 + birth1$numbirth_gtc_whitenh_1519 + birth1$numbirth_somecoll_whitenh_1519 + birth1$numbirth_hsgrad_whitenh_1519 + birth1$numbirth_ba_whitenh_1519

# getting total number of births for black people from 15 to 19 years old  
birth1$numbirthblack15 = birth1$numbirth_lesshs_blacknh_1519 + birth1$numbirth_educ_miss_blacknh_1519 + birth1$numbirth_gtc_blacknh_1519 + birth1$numbirth_somecoll_blacknh_1519 + birth1$numbirth_hsgrad_blacknh_1519 + birth1$numbirth_ba_blacknh_1519

# getting total number of births for hispanics from 15 to 19 years old  
birth1$numbirthhisp15 = birth1$numbirth_lesshs_hisp_1519 + birth1$numbirth_educ_miss_hisp_1519 + birth1$numbirth_gtc_hisp_1519 + birth1$numbirth_somecoll_hisp_1519 + birth1$numbirth_hsgrad_hisp_1519 + birth1$numbirth_ba_hisp_1519

# getting total number of births for white people from 20 to 24 years old
birth1$numbirthwhite20 = birth1$numbirth_lesshs_whitenh_2024 + birth1$numbirth_educ_miss_whitenh_2024 + birth1$numbirth_gtc_whitenh_2024 + birth1$numbirth_somecoll_whitenh_2024 + birth1$numbirth_hsgrad_whitenh_2024 + birth1$numbirth_ba_whitenh_2024

# getting total number of births for black people from 20 to 24 years old
birth1$numbirthblack20 = birth1$numbirth_lesshs_blacknh_2024 + birth1$numbirth_educ_miss_blacknh_2024 + birth1$numbirth_gtc_blacknh_2024 + birth1$numbirth_somecoll_blacknh_2024 + birth1$numbirth_hsgrad_blacknh_2024 + birth1$numbirth_ba_blacknh_2024

# getting total number of births for hispanics from 20 to 24 years old  
birth1$numbirthhisp20 = birth1$numbirth_lesshs_hisp_2024 + birth1$numbirth_educ_miss_hisp_2024 + birth1$numbirth_gtc_hisp_2024 + birth1$numbirth_somecoll_hisp_2024 + birth1$numbirth_hsgrad_hisp_2024 + birth1$numbirth_ba_hisp_2024

# gets the total number of births by year, compiled nationally
birthhist = birth1 %>%
  filter(year >= 2000) %>%
  group_by(year) %>% 
  select(year,pop1544, numbirthwhite15,numbirthwhite20,numbirthhisp15,numbirthhisp20,numbirthblack15,numbirthblack20) %>%
  summarize(sumpop = sum(pop1544),white15 = sum(numbirthwhite15), white20 = sum(numbirthwhite20), black15 = sum(numbirthblack15), black20 = sum(numbirthblack20), hisp15 = sum(numbirthhisp15), hisp20 = sum(numbirthhisp20))

# gets the birth rate for each race by adding the number of births from the 15 to 19for each race, then multiplying by 10,000 and dividing by the female population
fig21<-birthhist %>%
  mutate(whiteall = white15*10000/sumpop, blackall= black15*10000/sumpop, hispall = hisp15*10000/sumpop) %>% 
  select(year, whiteall, blackall, hispall) %>%
  pivot_longer(cols = c("whiteall", "blackall", "hispall"),
               names_to = "racegroup",
               values_to = "value"
  ) %>%
  ggplot(mapping = aes(x = year, y = value, color = racegroup)) +
  geom_line() +
  theme_minimal() +
  scale_x_continuous(breaks = seq(2000,2020, by=4))+
  labs(title = "Age Group 15 to 19",
       x = "Year",
       y = "Birth Rate",
       color = "Race") +
  scale_color_brewer(palette = "Set1", labels = c("Black", "Hispanic", "White")) +
  theme(legend.position = "bottom")

fig22<-birthhist %>%
  mutate(whiteall = white20*10000/sumpop, blackall= black20*10000/sumpop, hispall = hisp20*10000/sumpop) %>% 
  select(year, whiteall, blackall, hispall) %>%
  pivot_longer(cols = c("whiteall", "blackall", "hispall"),
               names_to = "racegroup",
               values_to = "value"
  ) %>%
  ggplot(mapping = aes(x = year, y = value, color = racegroup)) +
  geom_line() +
  theme_minimal() +
  scale_x_continuous(breaks = seq(2000,2020, by=4))+
  labs(title="Age Group 20 to 24",
       x = "Year",
       y = "Birth Rate",
       color = "Race",
       caption = "Source: Birth Rates collected from
                  CDC Vital Statistics Births Reports
                  for 2015, 2019 and 2020.") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "none")

grid.arrange(fig21,fig22,ncol=2,nrow=1)

```

@fig-2 shows the trend in birth rates over the period 2000-2019 in different races for two different age groups. In all race group, the birth rates starts to drop significantly after the Great Recession. In a age group of 15-19, all three race groups had similar drops on the birth rate. On the other hand in age group 20-24 the white race group showed significanat birth rate drop from 86 to 49 birth per 1000 population., where as other race groups had much less birthrate drop.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-4
#| fig-cap: Changes in Other Economic factors and Birth Rate over the periods of 2004-2008 and 2014-2019

#04~19 birthrate data
birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))
pop1 = read_dta(here::here("inputs/data/pop/age_race_comp_seer.dta"))
birth1 <- birth1 %>% 
  filter((year>=2004 & year<=2008) | (year>=2015 & year<=2019))

birth1 <- birth1 %>% 
  subset(select=c(stname,year,numbirth1544))

pop1<-pop1 %>% 
  mutate(pop1544 = pop1519 + pop2034 + pop3544)%>% 
  filter((year>=2004 & year<=2008) | (year>=2015 & year<=2019)) %>%
  group_by(year,stname) %>%
  select(pop1544) %>% 
  summarise(across(everything(),sum))

fig4_data <- right_join(birth1, pop1, by=c("stname","year"))
fig4_data <- fig4_data %>% 
  mutate(year = case_when(year=2004 & year<=2008 ~ 2004,
                           year>=2015 & year<=2019 ~ 2019)) %>%
  group_by(year,stname) %>%
  select(numbirth1544,pop1544) %>% 
  summarise(across(everything(),sum)) %>% 
  mutate(birthrate=numbirth1544/pop1544*1000) %>% 
  select(-c(numbirth1544,pop1544))

fig4_data <- fig4_data %>% 
  group_by(stname) %>% 
  mutate(birthrate = birthrate[year == 2019] - birthrate[year == 2004])

#04~19 childcare cost for fig4b
childcare_aesc=read_dta(here::here("inputs/data/long_term/childcare_asec_09_1519.dta"))
childcare_aesc <- childcare_aesc %>% 
  mutate(childcare_cost=child_asec)
childcare_aesc <- childcare_aesc %>% 
  select(year,stname,childcare_cost)
fig4_data <- right_join(fig4_data, childcare_aesc, by=c("stname","year"))
fig4_data <- fig4_data %>% 
  group_by(stname) %>% 
  mutate(childcare_cost = childcare_cost[year == 2019] - childcare_cost[year == 2004])

#04~19 rentcost data for fig4c
rentcost_0419=read_dta(here::here("inputs/data/long_term/longdiff-rentcost-0419.dta"))

rentcost_0419 <- rentcost_0419 %>% 
  mutate(stname = fips(rentcost_0419$statefip, to='Abbreviation')) %>%
  group_by(stname) %>% 
  mutate(avg_grossrent = avg_grossrent[year == 2019] - avg_grossrent[year == 2004]) %>% 
  select(year,stname,avg_grossrent)

fig4_data <- right_join(fig4_data, rentcost_0419, by=c("stname","year"))

#04~19 wage ratio data for fig4d
rhs_1519=read_dta(here::here("inputs/data/long_term/longdiff-RHS-1519.dta"))
rhs_0408=read_dta(here::here("inputs/data/long_term/longdiff-RHS-0408.dta"))
rhs<-rbind(rhs_1519, rhs_0408)
rhs <- rhs %>% 
  mutate(stname = fips(rhs$statefip, to='Abbreviation')) %>% 
  mutate(fem_male_ratio=p50wage_all/p50wage_male_all) %>% 
  select(year,stname,fem_male_ratio) %>% 
  group_by(stname) %>% 
  mutate(fem_male_ratio = fem_male_ratio[year == 2019] - fem_male_ratio[year == 2004])

fig4_data <- right_join(fig4_data, rhs, by=c("stname","year"))

#04~19 student loan data for fig4e
std_loan=read_dta(here::here("inputs/data/long_term/nyfed-student-debt-per-capita-0419.dta"))

std_loan<- std_loan %>% 
  group_by(stname) %>% 
  mutate(stud_loan_cap = stud_loan_cap[year == 2019] - stud_loan_cap[year == 2004])

fig4_data <- right_join(fig4_data, std_loan, by=c("stname","year"))

fig4_data <- fig4_data %>% 
  head(51) %>% 
  select(-c(year))

#drawing
fig4b<-fig4_data %>% 
  ggplot(aes(x=childcare_cost*1000, y=birthrate)) +
  geom_point(size=0)+
  labs(
    title = "Childcare Expenditures",
    x = "US dollar",
    y = "Birthrate") +
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(0, 5000), breaks=seq(0, 5000, by = 1000),labels = paste0("$",seq(0, 5000, by=1000))) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

fig4c <-fig4_data %>% 
  ggplot(aes(x=avg_grossrent, y=birthrate)) +
  geom_point(size=0)  +
  labs(
    title = "Rental Housing Costs",
    x = "US dollar",
    y = "Birthrate") +
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(0, 400), breaks=seq(0, 400, by = 75),labels = paste0("$",seq(0, 400, by = 75))) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

fig4d<-fig4_data %>% 
  ggplot(aes(x=fem_male_ratio, y=birthrate)) +
  geom_point(size=0)  +
  labs(
    title = "Female-Male Wage Ratio",
    x = "Fem to Male Wage Ratio",
    y = "Birthrate")+
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(-0.08, 0.12), breaks=seq(-0.08, 0.12, by = 0.04)) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

fig4e<-fig4_data %>% 
  ggplot(aes(x=stud_loan_cap*1000, y=birthrate)) +
  geom_point(size=0)  +
  labs(
    title = "Student Loan Debt",
    x = "US dollar",
    y = "Birthrate") +
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(1000, 4500), breaks=seq(1000, 4500, by = 750),labels = paste0("$",seq(1000, 4500, by=750))) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

grid.arrange(fig4b,fig4c,fig4d,fig4e,ncol=2,nrow=2)

```

@fig-4 shows an interesting relationship between changes in birthrate and other economic factors over five years period. Each marker represents a state labelled with its two-letter abbreviation.

Annual expenditures on childcare for families with children under 12 who report positive childcare spending rose \$2,170 nationwide for the average of 2015--2019, based on data from the Current Population Survey (all dollar values throughout are measured in constant 2019 dollars). But in the cross-state correlations, places where childcare expenditures increased more, did not experience a noticeable drop in birth rates. In fact, the trend shows that more increases in childcare expenditure have no significant impact on birthrate changes.

Over this period, average monthly rents for a two- to three-bedroom apartment rose \$124 per month (from \$930 to \$1,060 in 2018 dollars, a 14 percent increase) nationwide. The increase was much larger in some states, like Colorado, Washington, and the District of Columbia. The data, however, do not indicate a negative association between rent changes and birth rate changes.

We used the female-to-male median earnings ratio among full-time, year-round workers to examine women's economic opportunities. On average, the median earning ratio rose to 0.04 in the period. 12 states where the wage ratio rose the most over this sample period did not show a greater birth rate decline.

If young adults are saddled with debt, they might not have sufficient disposable income to have a child or more children. We consider the total student debt per capita level in a state, which has increased from \$2,500 to \$5,400 (in 2018 dollars) on average. The relationship between state-level student debt and the birth rate is generally flat. It indicates that increases in student debt are related to aggregate reductions in birth rates.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-5
#| fig-cap: Trends in Birth Rates by Age Groups from 1990-2019

# Reading in the data.

birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))

birth1 <- birth1 %>%
  group_by(year) %>%
  select(-c(stname)) %>% 
  summarise(across(everything(),sum))

birth1 <- birth1 %>% 
  mutate(birthrate1519 = numbirth1519/pop1519*1000) %>% 
  mutate(birthrate2024 = numbirth2024/pop2024*1000) %>% 
  mutate(birthrate2529 = numbirth2529/(pop25+pop26+pop27+pop28+pop29)*1000) %>% 
  mutate(birthrate3034 = numbirth3034/(pop30+pop31+pop32+pop33+pop34)*1000) %>% 
  mutate(birthrate3539 = numbirth3539/(pop35+pop36+pop37+pop38+pop39)*1000) %>% 
  mutate(birthrate4044 = numbirth4044/(pop40+pop41+pop42+pop43+pop44)*1000)

birth1 <- birth1 %>% 
  select(year, birthrate1519, birthrate2024, birthrate2529, birthrate3034, birthrate3539, birthrate4044) %>%
  pivot_longer(-c(year),
               names_to = "agegroup",
               values_to = "birthrate"
  )

birth1<-birth1 %>% 
  mutate(agegroup = as.character(str_replace(agegroup, "birthrate", "")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "1519", "15~19")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "2024", "20~24")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "2529", "25~29")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "3034", "30~34")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "3539", "35~39")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "4044", "40~44")))

birth1 %>% 
ggplot(mapping = aes(x = year, y = birthrate, color = agegroup)) + 
  geom_line()  +
  labs(
    title = "",
    x = "Year",
    y = "birthrate",
    caption = "Source: CDC Vital Statistics Births Reports.",
    color = "Age Group"
  ) +
  scale_x_continuous(limits=c(1991, 2020), breaks=seq(1991, 2020, by = 8)) +
  scale_y_continuous(limits = c(0,125), breaks = seq(0,125, by=20)) +
  theme_light()
```

@fig-5 shows lots of important information about when people tend to have kids at certain age over the timeframe from 1990 to 2020. Based on the graph, the data shows that more people have kids after the age of 30 during this present time. In the past, most people tend to have kids before the age of 30. Around 2000 to 2005, the lifestyle started to change, where people tended to either get married or have kids later, which is why the trend of birth rates in age 30 and up has been going higher since the 2000s. While births before the age of 30 tend to slow down and decrease over time from the 2000s. According to the graph, maybe in another 100 years, we can expect that most birth rates can happen around 40-44.

There are drastic changes shown in @fig-5, such as age 30-44 tends to increase around the 2000s, while the ages below 30 decreases for the number of birth rates. Around the decade of the 2000s or the beginning of this new 21st century, there was lots of modernization from people's lifestyle to other things on Earth, which is coming from several factors such as education and employment due to new innovations and inventions [@citedimock]. There are lots of new behaviour being introduced to a new generation of people. So, as they get new behaviour of attitudes, that is being involved in their decisions of marriages, and when to have kids. This is why lots of people are now getting married late and/or having kids much later when compared to the past.

Figure 5. Children Ever Born by Mother's Age and Birth Cohort (@fig-6).

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-6
#| fig-cap: Children Ever Born by Mother's Age and Birth Cohort

nch_cohot_data <- read_dta(here::here("inputs/data/nchs/nchs_cohort_analysis.dta"))
agecomp_data <- read_dta(here::here("inputs/data/pop/agecomp-seer.dta"))

pop_by_cohort<- agecomp_data %>%
  select(1:32) %>%
  group_by(year) %>%
  select(-c(stname)) %>%
  summarise(across(everything(),sum)) %>%
  pivot_longer(-c(year), names_to = "mage", values_to = "pop") %>%
  mutate(mage = as.numeric(str_replace(mage, "fem", ""))) %>%
  mutate(birthyear = year - mage) %>%
  mutate(cohort2 = case_when(birthyear >= 1968 & birthyear <=1972 ~ 1,
                            birthyear >= 1973 & birthyear <=1977 ~ 2,
                            birthyear >= 1978 & birthyear <=1982 ~ 3,
                            birthyear >= 1983 & birthyear <=1987 ~ 4,
                            birthyear >= 1988 & birthyear <=1992 ~ 5,
                            birthyear >= 1993 & birthyear <=1997 ~ 6)) %>% 
  group_by(cohort2,mage) %>%
  summarise(pop = sum(pop))

fig5_data <- right_join(nch_cohot_data, pop_by_cohort, by=c("mage", "cohort2"))
fig5_data <- fig5_data %>%
  mutate(birthrate=numbirth/pop*1000,
         age_2024_year = case_when(cohort2 == 1 ~ 1992,
                                   cohort2 == 2 ~ 1997,
                                   cohort2 == 3 ~ 2002,
                                   cohort2 == 4 ~ 2007,
                                   cohort2 == 5 ~ 2012,
                                   cohort2 == 6 ~ 2017))

fig5_data <- fig5_data %>% 
  group_by(age_2024_year) %>% 
  mutate(cum_birthrate = cumsum(birthrate)/1000, 
         cohort = case_when(cohort2 == 1 ~ "1968-1972        1992",
                                  cohort2 == 2 ~ "1973-1977        1997",
                                  cohort2 == 3 ~ "1978-1982        2002",
                                  cohort2 == 4 ~ "1983-1987        2007",
                                  cohort2 == 5 ~ "1988-1992        2012",
                                  cohort2 == 6 ~ "1993-1997        2017"))

fig5_data <-  na.omit(fig5_data)
fig5_data %>%
  ggplot(aes(x = mage,
             y = cum_birthrate,
             color = cohort)) +
  geom_line()  +
  labs(
    title = "",
    x = "Ages",
    y = "Cumulative birthrate",
    color = "Birth cohorts   Age 20 in"
  ) +
  scale_x_continuous(limits=c(15, 44), breaks=seq(15, 44, by = 3)) +
  theme_minimal()+
  theme(legend.position = "right")

```

@fig-6 shows lots of useful information about the cumulative birth rate by mother's age in several birth cohorts from early 1968 to late 1997. We tend to use these years since, in recent times, we believe that there is enough data for these years since most of the births and living people are among these cohorts. According to @fig-6, the earlier the cohort is, the cumulative birth rate will be really high. While the later the cohort is, the cumulative birth rate will be much lower. The highest cumulative birth rate is coming from the birth cohort of 1968-1972 and 1973-1977. While the lowest comes from the 1993-1997 birth cohort.
\newpage

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-1
#| tbl-cap: The Eight Demographic Groups that Contributed the Most to the 2007-2019 Decline in the U.S. Brith Rate
numbirth_2044 <- read.csv(here::here("inputs/data/edu/numbirths_educ_2044.csv"))
numbirth_2044 %>% 
  kable(col.name=c("Age Group","Change in Birth Rate"),
        booktabs = TRUE,show_col_types=FALSE,linesep="",
        title = "The Eight Demographic Groups that Contributed to the Most to the 2007-2019 Decline in the U.S. Brith Rate"
        )

```

@tbl-1 shows the most contributor to the decline in birth rates across all age groups. The top contributor is Hispanic with less than high school education women between age 20-24 at -88.3, Hispanic with high school education women between age 15-19 at -57.5, and Black (Non-Hispanic) women between age 15-19 at -39.8, respectively.

\newpage

# Discussion

## Birth Rates by Demographic Groups {#sec-first-point}

In 1997, NCHS released a report titled "Mother's Educational Level Influences Birth Rate." According to the report, a women's education level is the best predictor of how many children she will have. In the report, NCHS describes a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment [@citeNCHS].\
We are particularly interested in ethnic groups of ages 15-19 and 20-24. @fig-2 shows a sharp decline in birth rates among all groups of women between the age of 15 to 19 during the post-2007 recession. @fig-2 also shows a sharp decline over the same period; however, only among white ethnic women between the ages 20 to 24, while other groups show a relatively steady decline. Figures that are omitted from this paper of the other age groups (25 to 34 and 35 to 44) show a similar trend both in pre and post-recession. However, as shown in @tbl-1, the most significant contributor to the decline in the birth rate is Hispanic, with less than High School education, women between ages 20-24 after 2007. This result contradicts the earlier claim made by NCHS. The finding suggests, in the post-2007 recession, the women's education level and birth rate now correlate positively.

## Economic and Policy Factors {#sec-second-point}

In @fig-4, we show other economic factors that might have negatively affected the birth rate in the post-2007 recession across the states. Although the graphs show their inverse relationship, no dramatic change is shown. For instance, as there were somewhere around \$1000 to \$2000 increases in childcare expenses over the past 13 years, we see that the decrease in the actual birth rate is only between 0 to -10. Change in per capita student loans also has increased between \$2500 to \$3250, but it only contributed to the decrease somewhere between 0 to -5 across the states. Based on these findings, it is hard to find an explanation of the dramatic decline in the birth rate since 2007 from economic factors.

## A Potential Explanation: Shifting Priorities {#sec-third-point}

When it comes to the age of the population of the United States and comparing it from the past to now. It shows that more people tend to have kids much later than before. While fewer people have kids earlier than in the past. In the past, more people tend to have kids much earlier. There are several reasons why more people had kids at an early age back then; people used to get married much earlier, and many people were unemployed or doing a job for self-employment, which was not even an educated or professional job. The modern lifestyle that has been adopted in the past few decades tends to make people focus on all other areas of life, which makes them marry much later, and have kids later. Nowadays, many marriages happen at the age of late 20s or early 30s, while they tend to have their first child around the early to mid-30s. @fig-5 depicts the birth rates in women between the ages 30-39 and 35-39 positive slope trend. Whether pre or post-2007 recession, this trend looks steady over the past 40 years. This finding aligns with the current common belief and the social norm where, perhaps, women's peak childbearing ages have shifted from the early twenties to early thirties. @fig-6 shows the average number of children born to women by age in five-year intervals from the 1968-1972 through 1993-1997 birth cohorts. These birth cohorts include women entering their peak childbearing years (20--24) in 1992, 1997, 2002, 2007, and 2017. This chart indicates that the three cohorts of young adult women born between 1968 and 1982 in 1992, 1997, and 2002 had similar childbearing age profiles. The 2007 birth cohort of women had fewer children in their 20s and 30s. The 2012 and 2017 birth cohorts (born between 1988 and 1997) have significantly fewer children than the preceding cohorts. This finding is consistent with a decline in births, primarily to more recent cohorts of women. From a statistical point of view, this divergence across cohorts explains the sizable decrease in annual birth rates that began in 2007.

## Weakness and Moving Forward

### Bias and Ethical Concerns

As we know that this data was collected and created by other human beings, just as how we are. There are some ethical concerns that could affect the accuracy of our data because humans can make mistakes, and they can make decisions based on their preferences, which can affect the whole dataset. So, this analysis is only valid based on our dataset, which can be modified if the original dataset is changed.

The original data shows different races, such as White, Black, and Hispanic. They have included Hispanic based on all colours, which erases the identity of Black and Indigenous Hispanic people. This tends to be a bit of racism because the information provided on the data did not specifically indicate whether or not the Hispanic category includes everyone that identifies as Hispanic or is it just people that look Hispanic.

### Data Collection Concerns

This dataset tends to collect huge data information on the population among people from the United States and considers lots of factors from race, birth rates by age, student debts, etc. Having too much information means that when the data is collected, there is a possibility of human error, calculation error, and chances of small modification of data due format of reporting an error. So, the way how the data was collected can have lots of problems because people can make mistakes easily.

### Moving Forward

This paper raised concerns about a dramatic decline in U.S. birth rates starting in 2007. Through the paper, we proposed a general explanation for the decline in births across recent cohorts of U.S. women that focuses on the shifting priorities of cohorts. As dictated by the recent socio-phenomena that encompasses preferences for having children, life aspirations, and the nature of parenting, a change in the relationship between education and childbearing, a rise in childlessness, and the establishment of a two-child norm for those having children have irreversibly changed [@citepaper]. This new social phenomenon that might be causing the dramatic decline in U.S. birth rates is important from an economic policy standpoint to acknowledge that an aging population and shrinking workforce pose challenges to economic growth and the sustainability of social insurance systems [@Yang2003]. Thus, the most appropriate way to address declining U.S. birth rates may be required: 1. developing A.I. and robots where a shrinking workforce poses a problem, 2. an appropriate policy to address issues with Social Security and Medicare on a secure basis for the long-term. Suppose the U.S. birth rate is not showing any reverse trend shortly. In that case, the U.S. economy and political system should enforce greater emphasis on addressing the issue before it becomes irreversible.

\newpage

# References